---
title: "Evolution of cinema's frequentation in France"
author: "Kylian SANTOS and Cédric BAZILE"
output:
  html_document:
    toc: TRUE
editor: visual
---

<center>

# Github link

</center>

<center><https://github.com/KylianSts/CNC_INSEE_project></center>

```{r Configuration, warning=FALSE, message = FALSE}
here::i_am("CNC_INSEE_project.Rproj")
library(here)
library(dplyr)
library(tidyr)
library(readxl)
library(readr)
```

# Formating and cleaning of the CNC database

<center>

# A presentation of the project

</center>

## Sources

This project is powered by two databases :

-   CNC's "**Exploitation - data by department**" : <https://www.cnc.fr/cinema/etudes-et-rapports/exploitation--donnees-par-departement_223440>

-   INSEE's "***Harmonized population census data from 1968***" : <https://www.insee.fr/fr/statistiques/6671801?sommaire=2414232>

### "Exploitation - data by departement"

*Original french name : "**Exploitation - données par département**"*

The first database was created by the French National Center for Cinema and the Moving Image, most commonly known as CNC. CNC is an agency of the French Ministry of Culture. Its role is to implement the state's policy on the question of cinema, audiovisual and multimedia. It is notably in charge of the regulation, promotion and distribution of french movies. The French agency also collects data regarding its purview and proposes part of them in open data.

The database "Exploitation - data by department" is one of the aforementioned opan data. It is a statistical monitoring on the evolution of french cinema theater and its attendance between 1966 and 2022. It notably contains data on the number of venues, their characteristics (e.g. number of screens, seating capacity, ...); but also on the number of sold tickets and gained revenues. As its name indicates, french departments (according to their 2021 administrative division) are the statistical unit of the base. The database were constructed by using two main sources :

-   the **list of cinemas** authorized by the agency

-   and their "bordereaux des recettes" which can be roughly translated by "**revenue slips**"

### "***Harmonized population census data from 1968***"

*Original title : Données harmonisées des recensements de la population à partir de 1968*

The second database was produced by Insee, the french National Institute of Statistics and Economic Studies. Insee is French national bureau and one of its bigger data provider. The data that will be used for its project has for sources demographic information which come frome the different population censuses between 1968 and 2019.

```{r Summary_of_the_last_census, echo=FALSE, message = FALSE}
list_of_census <- data.frame(census = paste0("n°", 1:8),year=c(1968,1975,1982,1990,1999,2008,2013,2019))

list_of_census<- list_of_census |>  pivot_wider(names_from="census",values_from = "year")

list_of_census |> knitr::kable(caption ="French census between 1968 and 2019", format="html", align='c')
```

Censuses are one of the most important data collections taking place in France. Answering their survey is an obligation for all inhabitants under penalty of a fine (article 3 of the Law n°51-711 of 7 June 1951). Historically, population censuses took the form of an exhaustive collect of all French townships during asame year. However, the collection method evolved in 2002. Nowadays, Insee uses a rotational approach. For townships of at least 10 000 inhabitants, 8% of the housing of are interviewed each year. The goal is then to cover 40% of their population over a 5-year period. This 5-year moving sample is then used to estimate demographic data annually. For townships withless of 10 000 inhabitants, the collection is still exhaustive but is not simultaneous. Indeed, it occurs in a fifth of towns each year ensuring that all townships are censused every 5 years. Demographic information between 2 collection is then estimated using data from the nearest one.

The data is collected by face-to-face interviews or-since 2015- via internet surveys.

The database that will be used in this project is an harmonized compilation of data from 8 census databases. Its content includes the number of inhabitant by department and key socio-demographic variables (e.g. age, socio-professional categories, type of professional activity).

It can be downloaded from the [INSEE website](https://www.insee.fr/fr/statistiques/6671801?sommaire=2414232) in a DBF or CSV format.

### Modification of the original databases

The two databases underwent transformation to be exploitable for our project. The main problem of the Insee one is that it consumes a lot of RAM because of its size (3GB). Indeed some categorial variables have an important number of "categories"/"levels". For example, the variable on the age of the population contains 120 levels, one by each possible year of life of an individual. The result is that the number of potential combinations, and therefore of rows, among all variables is impressively large contributing to the important size of the database.

```{r Computation_of_the_number_of_rows_in_Insee_cleanv3, echo=FALSE, message = FALSE}

Insee_cleanv3 <- read_csv("INSEE_cleanv3.csv")

In_nrow <- Insee_cleanv3 |> nrow()
```

Our work was to reduce the number of these combinations in order to downsize the database. To do so, we transform the 120 levels of the variable corresponding to the age into 5 age ranges. Furthermore, we get rid of the variables that were not useful to our project using a group_by function. It allows us to go from a database of more than 50 millions of rows to a CSV of `r In_nrow` rows.

The process that we used is detailed in the quarto document entitled "Parallel_code_Insee" that is available in this R project. We ha ve preferred to put it in a separate qmd because of the important size of the database linked with its code.

The CNC's data was less memory-consuming. However their format was not convenient for coding with R. Indeed, the original XSLX document put each variable in an independent sheet. In the codes below, we merge all this independent sheets into one uniform database.

#### Formating and cleaning of the CNC database

```{r Cleaning_CNC_database, message = FALSE}

file_path <- "exploitation_données_par_département.xlsx"

cnc_data <- read_excel(file_path, sheet = 4, skip = 5)

cnc_data_long <-  cnc_data |>
  select(-colnames(cnc_data)[2]) |>
  rename(departement_code = colnames(cnc_data)[1]) |>
  pivot_longer(
    cols = -departement_code,
    names_to = "year", 
    values_to = excel_sheets(file_path)[4]
    )

cnc_data_merge <- cnc_data_long

for (i in 5:which(excel_sheets(file_path) == "tmofAE")) {
  cnc_data <- read_excel(file_path, sheet = i, skip = 5)

cnc_data_long <-  cnc_data |>
  select(-colnames(cnc_data)[2]) |>
  rename(departement_code = colnames(cnc_data)[1]) |>
  pivot_longer(
    cols = -departement_code,
    names_to = "year", 
    values_to = excel_sheets(file_path)[i]
    )

cnc_data_merge <- left_join(cnc_data_merge, cnc_data_long, by = c("departement_code", "year"))
}
```

As mentioned above, the original XSLX file has multiple sheets : each sheet being a variable with :

-   rows as departements

-   and columns as years.

To permit the merger between each sheets, the prime step was to rename the first column so that we make sure that each sheet shares the same name. Then the second column which is the department was dropped. It permit us to start the merger.

```{r Suppression_of_total_rows, message = FALSE}
cnc_data_clean <- cnc_data_merge |>
  filter(!is.na(departement_code))
```

Each sheet has a row wich represents the total. We preferred to drop it in order to simplify next statistics treatment.

```{r replace_NA_by_0_multiplexes, message = FALSE}
multiplexes_with_0 <- cnc_data_clean |> 
  select(multiplexes, year, departement_code) |> 
  pivot_wider(names_from = year, values_from = multiplexes) |> 
  mutate(across(where(~any(!is.na(.))), ~ifelse(is.na(.), 0, .))) |> 
  pivot_longer(cols = -departement_code, names_to = "year", values_to = "multiplexes")

cnc_data_clean <- cnc_data_clean |> select(-multiplexes) 
  
cnc_data_clean <- left_join(cnc_data_clean, multiplexes_with_0, by = c("departement_code", "year"))
```

In the original "Multiplexes" sheet, NA value were assigned when there were 0 multiplexes. We replaced this NA value with 0. and ensured that this replacement only occurs for the year when data had been collected

### Description of the final bases before merging

The final result is two modified datasets.

**cnc_data_clean**

```{r Nb_rows_and_columns_CNC, echo=FALSE, message = FALSE}
cnc_data_clean %>%
  summarize("Number of rows" = prettyNum(nrow(.), big.mark = " "),
            "Number of columns" = ncol(.)) %>%
  knitr::kable(caption = "Description of the database", format = "html", align = 'c')
```

**Insee_clean_v3**

```{r Nb_rows_and_columns_Insee, echo=FALSE}
summary_table <- Insee_cleanv3 %>%
  summarize("Number of rows" = prettyNum(nrow(.), big.mark = " "),
            "Number of columns" = ncol(.))

knitr::kable(summary_table, caption = "Description of the Insee_cleanv3 database", format = "html", align='c')
```

</center>

#### Description of the variables

</center>

We have in the CNC database different variables :

-   "**établissement**" represents the **number of movie theaters** for each department
-   "**écrans**" represents the **number of screens,** i.e. movie rooms for each department
-   "**fauteuils**" represents the total **number of seats** in all movie theaters for each department
-   "**multiplexes**" represents the **number of** **multiplexes** for each department . A multiplex is a movie theater complex with multiple screens within a single complex.
-   "**séances**" represents the total **number of seances [in thousand]{.underline}** for each department
-   "**entrées**" represents the total **number of tickets bought [in million]{.underline}** for each department
-   "**recettes**" represents the **total** **revenue in million** for each department
-   "**RME**" represents the **average revenue by bought tickets**
-   "**Taux d'occupation des fauteuils**" is the ratio between **entries during the year and the product between the number of chairs and the number of sessions**. Taux d'occupation des fauteuils makes it possible to assess the [**"fullness" of the rooms.**]{.underline}

And we have the same for Arthouse. the term Arthouse refers to so-called auteur films, economically independent, with small budgets, open to the different cultures of the world. An Arthouse cinema is a cinematographic establishment which exhibits a significant proportion of recommended Arthouse films and which supports these often difficult films with an adapted animation policy.

Insee database is a sociodemographic description of each department with variables such as :

-   "**DEP_RES_21**" represents the **department of residence where individuals live**
-   "**AGE_GROUP**" is our computed data representing the a**ge of the population by range of 15 years**.
-   "**TYP_ACT**" give us the **type of activity** .
-   "**CSP**" is a categorial data that has for levels **socio-professional category**
-   "**POND2**" gives us **the number of** **inhabitants** a**ccording the combination of all previous variables**

</center>

#### Merging the INSEE and CNC database

</center>

```{r insee_database}
insee_data <- read.csv("INSEE_cleanv3.csv")

insee_data <- insee_data |> rename(departement_code = DEP_RES_21, year =  AN_RECENS)

insee_data_clean <- insee_data |> filter(!is.na(AGE_GROUP))
```

```{r mergeing}
cnc_data_clean$year <- as.integer(cnc_data_clean$year) 

data_merge <- left_join(cnc_data_clean, insee_data, by = c("departement_code", "year"))

data_clean <- data_merge |>
  filter(!is.na(POND2))
```

```{r}
data_clean |> group_by(year) |>  summarise(across(everything(), ~sum(is.na(.)), .names = "missing_{.col}"))
```
