---
title: "Parallel_code_Insee"
author: "C. BAZILE & K. SANTOS"
format: html
editor: visual
---

## Presentation of the documents

The codes below describeb the stepped we used to reduce the size of the initial Insee's 3 Go database in a smaller - and therefore more workable - database.

The INSEE's database is intitled : "***Données harmonisées des recensements de la population à partir de 1968***".

It was not possible to put the database on Github but it can be donwload on the following link : <https://www.insee.fr/fr/statistiques/6671801?sommaire=2414232>

In English, it can be translate by : "***Harmonized population census data from 1968***"

It contains socio-demographic information on the french population by department between 1968 and 2019.

The format of the database make it quite memory-consuming.

To permit an easier exploitation, we work on a solution to reduce the size of the aforementioned database. This document reconstruct the steps of our work.

## Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
