---
title: "Parallel_code_Insee : A reduction of a 3 Go Insee's database"
author: "C. BAZILE & K. SANTOS"
format: html
editor: visual
execute :
  keep-md: TRUE
  warning: FALSE
---

# Presentation of the document

The codes below describeb the stepped we used to reduce the size of the initial Insee's 3 Go database in a smaller - and therefore more workable - database.

The INSEE's database is intitled : "***Données harmonisées des recensements de la population à partir de 1968***".

It was not possible to put the database on Github but it can be donwload on the following link : <https://www.insee.fr/fr/statistiques/6671801?sommaire=2414232>

In English, it can be translate by : "***Harmonized population census data from 1968***"

It contains socio-demographic information on the french population by department between 1968 and 2019.

The format of the database make it quite memory-consuming.

To permit an easier exploitation, we work on a solution to reduce the size of the aforementioned database. This document reconstruct the steps of our work.

# Codes

## Setup

```{r}
 here::i_am("CNC_INSEE_project.Rproj")

library(data.table) 

# The  "data.table" package will be  one of our keytools. Indeed, its characteristics allows us to exploit easely big database what will be usefull in our situation 

library(here)
library(dplyr) library(tidyr) library(kableExtra)
```

## Importation of the data using the data.table pakcage
To reproduce this step download the database on : <https://www.insee.fr/fr/statistiques/6671801?sommaire=2414232>

```{r}

insee <- fread("Ficdep19.csv")
```

